Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2014-11-26 manuscript.tex
Correct distance and size formula
They now reflect the distance to the center of the tesseroid. Fixed the
number of divisions for the division example.

In our implementation,
we measure the distance between
the computation point
and the center of the top face tesseroid
(see Figure~\ref{fig:division}a).

The distance $d$ is

\begin{equation}
    d = \left[
        r^2 + r_2^2 - 2 r r_2
        \cos(\sin\phi\sin\Delta\phi + \cos\phi\cos\Delta\phi
            \cos(\lambda - \Delta\lambda))
        \right]^{\frac{1}{2}},
    \label{eq:distance}
\end{equation}

UPDATE THE EQUATION TO THE CENTER OF THE TESSEROID + NUMBER OF DIVISION IN
EXAMPLE FIGURE.

where
$(r, \lambda, \phi)$ are the coordinates of the computation point,
$\Delta\lambda = 0.5(\lambda_2 + \lambda_1)$,
and
$\Delta\phi = 0.5(\phi_2 + \phi_1)$.

We define the dimensions of the tesseroid
along longitude, latitude, and radius, respectively, as

\begin{equation}
    L_\lambda = r_2 \arccos(\sin^2\Delta\phi +
        \cos^2\Delta\phi\cos(\lambda_2 - \lambda_1)),
    \label{eq:sizelon}
\end{equation}
\begin{equation}
    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
\end{equation}
\begin{equation}
    L_r = r_2 - r_2.
    \label{eq:sizer}
\end{equation}

USE THE MAX OF SOUTH AND NORTH FOR DLON? MIGHT MAKE IT BETTER IN THE POLES.

@@

For $D = 1$ (Figure~\ref{fig:division}b) the number of tesseroids in the
adaptive discretization is 12,
for $D = 2$ (Figure~\ref{fig:division}c) it is 52,
and for $D = 6$ (Figure~\ref{fig:division}d) it is 1092.

\begin{figure}
    \centering
    \includegraphics{figs/tesseroid-split}
    \caption{
        Adaptive discretization
        of the tesseroid shown in (a)
        for a computation point P
        using the distance-size ratio $D$ equal to
        (b) 1, (c) 2, and (d) 6.
        $d$ is the Cartesian distance between
        the center of the top face of the tesseroid
        and $P$.
        $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
        Note that increasing $D$
        results in a fine division of the tesseroid
        close the computation point
        and a coarser division further away.
        The number of tesseroids in the discretization is (b) 12, (c) 52,
        and (d) 1092.
    }
    \label{fig:division}
\end{figure}

@@

\subsection{Computational performance as a function of $D$}

For close P, vary $D$ and graph computation time.

>>>>

Instead of the smallest distance,
we use in our implementation
the distance $d$ between
the computation point $(r, \lambda, \phi)$
and the geometric center of the tesseroid
$(r_t, \lambda_t, \phi_t)$

\begin{equation}
    d = \left[
        r^2 + r_t^2 - 2 r r_t
        \cos(\sin\phi\sin\phi_t + \cos\phi\cos\phi_t
            \cos(\lambda - \lambda_t))
        \right]^{\frac{1}{2}}.
    \label{eq:distance}
\end{equation}

We define the dimensions of the tesseroid
along longitude, latitude, and radius, respectively, as
(see Figure~\ref{fig:division}a)

\begin{equation}
    L_\lambda = r_2 \arccos(\sin^2\phi_t +
        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
    \label{eq:sizelon}
\end{equation}
\begin{equation}
    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
\end{equation}
\begin{equation}
    L_r = r_2 - r_2.
    \label{eq:sizer}
\end{equation}

@@

For $D = 1$ (Figure~\ref{fig:division}b) the number of tesseroids in the
adaptive discretization is 4,
for $D = 2$ (Figure~\ref{fig:division}c) it is 38,
and for $D = 6$ (Figure~\ref{fig:division}d) it is 936.

\begin{figure}
    \centering
    \includegraphics{figs/tesseroid-split}
    \caption{
        Adaptive discretization
        of the tesseroid shown in (a)
        for a computation point P
        using the distance-size ratio $D$ equal to
        (b) 1, (c) 2, and (d) 6.
        $d$ is the Cartesian distance between
        the center of the top face of the tesseroid
        and $P$.
        $L_r$, $L_\phi$, and $L_\lambda$ are the dimensions of the tesseroid.
        Note that increasing $D$
        results in a fine division of the tesseroid
        close the computation point
        and a coarser division further away.
    }
    \label{fig:division}
\end{figure}

@@
